### 2021-02-09 dags_for_scott.Rmd
fixed material

---
title: "Dags for Scott"
author: "Joseph Bulbulia"
date: "28/01/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
### Dag shoes two adjustment sets
```{r}
library(ggdag)
library(dagitty)
library(ggplot2)
library(tidyverse)

# more complex
dg <- dagify(
  ms  ~  mg + time,
  mg ~ time + L,
  L ~ time, 
  exposure =  "mg",
  outcome =   "ms"
) %>%
  tidy_dagitty()

colliders<-ggdag::ggdag_collider(dg,
                      from = "mg",
                      to = "ms",
                      control_for(c("time", "L")))+   theme_dag_blank() + labs(title = "colliders")
colliders
p1<-ggdag(dg)+ theme_dag_blank() + labs(title = "dag")

# graph adjustment sets
p2 <-ggdag::ggdag_adjustment_set(dg) + labs(title = "adjustment set") +
  theme(legend.position = "bottom") + theme_dag_blank() 
p1 + p2 + plot_annotation(title = "To infer path from MG --> MS we only need to condition on time",
                          subtitle = "assume:  
                          ms ~ mg + time, 
                          mg ~ time + L, 
                          L ~ time, 
                          exposure =  mg,
                          outcome =   ms")

ggdag_dseparated(dg,
  from = "mg",
  to = "ms",
  controlling_for = c("time", "L"),
  collider_lines = TRUE
) + theme_dag_blank() + labs(title = "d-relationships")

p1 + plot_annotation(title = "To infer path from MG --> MS we only need to condition on time")

# ggdag_drelationship(dg,
#   from = "mg",
#   to = "ms",
#   controlling_for = c("time","L"),
#   collider_lines = TRUE
# ) + theme_dag_blank()

##play
dg2 <- dagify(
  mg  ~ time + L,
  ms ~  time + L,
  mg ~~ ms,
  L ~ time, 
  exposure =  "ms",
  outcome =   "mg"
) %>%
  tidy_dagitty()

ggdag::ggdag_collider(dg2,
                      from = "ms",
                      to = "mg",
                      control_for(c("time", "L"))) +   theme_dag_blank()

# graph adjustment sets
ggdag::ggdag_adjustment_set(dg2, node_size = 14) +
  theme(legend.position = "bottom") + theme_dag_blank()

## SIMPLE
dg2 <- dagify(
  mgPRE  ~ time,
  ms ~  mgPRE + time + U,
  U ~ time, 
  exposure =  "mgPRE",
  outcome =   "ms"
) %>%
  tidy_dagitty()

jb1<-ggdag::ggdag_collider(dg2,
                      from = "mgPRE",
                      to = "ms",
                      control_for(c("time", "U"))) +   theme_dag_blank() + labs(title = "colliders")
jb1

# graph adjustment sets
jb2<- ggdag::ggdag_adjustment_set(dg2, node_size = 14) +
  theme(legend.position = "bottom") + theme_dag_blank()  + labs(title = "adjustment set")

jb2
library(patchwork)
jb1 / jb2 + plot_annotation(title = "If we think mega-societies and moralising gods co-evolve then we we might not get an obtain an unbaised estimate", tag_levels = "a")

#####

# more complex
tidy_ggdag2 <- dagify(
  wb  ~ rel + age + male + ses + cntry + den,
  rel ~ age + male + cntry + den,
  ses ~ age  + cntry,
  cntry ~ lang + den, 
  exposure =  "rel",
  outcome =   "wb"
) %>%
  tidy_dagitty()

#Here's the trouble with Scott's model
ggdag::ggdag_collider(tidy_ggdag2,
                      from = "rel",
                      to = "wb",
                      control_for(c("ses", "male", "cntry", "age","den","lang"))) +   theme_dag_blank()

# graph adjustment sets
ggdag::ggdag_adjustment_set(tidy_ggdag2, node_size = 14) +
  theme(legend.position = "bottom")

ggdag_drelationship(tidy_ggdag2,
  from = "rel",
  to = "wb",
  controlling_for = c("ses", "male", "cntry", "age", "den","lang"),
  collider_lines = TRUE
) + theme_dag_blank()

```

```{r}
tidy_ggdag2 <- dagify(
  wb  ~ rel + age + male + ses + cntry + den,
  rel ~ age + male + cntry + den,
  ses ~ age  + cntry,
  cntry ~ den, 
  exposure =  "rel",
  outcome =   "wb"
) %>%
  tidy_dagitty()

#Here's the trouble with Scott's model
ggdag::ggdag_collider(tidy_ggdag2,
                      from = "rel",
                      to = "wb",
                      control_for(c("ses", "male", "cntry", "age","den","lang"))) +   theme_dag_blank()

# graph adjustment sets
ggdag::ggdag_adjustment_set(tidy_ggdag2, node_size = 14) +
  theme(legend.position = "bottom") 
```

```{r eval }
tidy_k6 <- dagify(
  k6  ~ covid + mean + rel + male + income + age + rest,
  mean ~ rel + male  + age + rest,
  rel ~ age  + male,
  income ~ edu, 
  exposure =  "rel",
  outcome =   "k6"
) %>%
  tidy_dagitty()

#Here's the trouble with Scott's model
ggdag::ggdag_collider(tidy_k6,
                      from = "covid",
                      to = "k6",
                      control_for(c("rel", "male", "mean", "age","income"))) +   theme_dag_blank()

# graph adjustment sets
ggdag::ggdag_adjustment_set(tidy_k6, node_size = 14) +
  theme(legend.position = "bottom")

ggdag::ggdag_collider(tidy_k6,
                      from = "covid",
                      to = "k6",
                      controlling_for = c( "male","age")) +   theme_dag_blank()

ggdag_drelationship(tidy_k6,
from = "covid",
                      to = "k6",
                      controlling_for = c( "male","age","covid"), collider_lines = TRUE
) + theme_dag_blank()

```

```{r eval = FALSE}
remotes::install_github("BS1125/CMAverse")
library(CMAverse)
cmdag(outcome = "K6", exposure = "cv", mediator = c("FT", "Work"), 
      basec = c("C1", "C2"), postc = NULL, node = FALSE, text_col = "black")

cmdag(
  outcome = "K6",
  exposure = "COV",
  mediator = "F",
  basec = "male",
  postc = "reld"
)

```

### Confounder bias if we include our theoretically proposed causal mediators!

Before proceeding with the model, note a curiosity. If our interest is in the direct effects of church attendance on well-being we should *not* include social support because this introduces a collider bias.