### 2022-02-06 response-reviewer.md
reorder repo

Reviewer: 1

We are grateful to the reviewer for taking the time to review our manuscript, which has  substantially improved it. Below we address all reviewer comments.

*The motivation of this article is great, and while its execution is reasonable given the constraints of the commentary format, I would encourage the authors to clean it up in a few key ways.*

- Thanks.

First, I should note that it comes across as hastily put together and lacking consideration for the reader.  For example, we should not have to read what appear to be unmodified xtable tables; we should be able to immediate interpret what the estimates are for in Tables B1-B4.

- Thank you. The tables have been placed in the supplementary materials. We do not provide coefficients to better control for the "table two" fallacy. Rational and citation for the table two fallacy is given.

Also, the figures are unreadable without being able to zoom in. The authors should assume readers will use paper to read their contribution. Other examples abound (see below).

- Thank you. We have increased the figure size to meet the journal's requirements, and attached the figures seperately.

I would also encourage the authors to expand this to a full-length article because it is clear that it requires a little more room to breathe.

- Thank you. We agree. However, this is not possible in for an RBB commentary.

That said, the commentary might benefit for some sharper targeting. The commentary should eventually get published, but it needs to guide the reader better: 1) tighten up the framing; 2) lend a bit more consideration to an audience you're trying to convince; your motivations are clear to me, but I doubt they would be to someone who doesn't already know them; 3) stitch the materials together a bit tighter; make sure each part is explicitly connected.

- Thank you. We agree. We have substantially revised the manuscript for clarity.

Framing: 
The MARP project was an attempt to crowd-source analyses, the authors are doing the project and social science more broadly a service by focusing on causal inference. I’m completely sympathetic with the approach and motivation (in fact, the charge is more applicable to the social sciences in general, the field of the scientific study of religion more specifically). The authors don’t imply that the MARP project is guilty of adopting a “model free” approach, but it *might* come across that way given how the framing is a little negative (“...illustrate how a ‘model free’ approach can mislead...”). This might be a little clearer if the scope of criticism were widened (i.e., to also apply to social science more generally).

- We have given a stronger argument against crowd-sourcing in the absence of expert knowledge for causal inference.

> QUOTE HERE

- We have included a reference to new reference to [Miguel Hernans] who makes this point in a recent publication.

Introduction: 
In a few ways, the introduction needs work (another example of its rushed style). First, it doesn’t obviously follow that “religion affects mental health” because it “remains a core feature of human life”. Second, the intro. enumerates three points that are neither introduced nor obviously related to one another. Initially, I thought it would help to say something along the lines of “Yet[,] inferring how religion affects mental health requires care [for at least three important reasons].” but I’m not sure if the following points are reasons for this care. Third, the second point requires some citations and/or precision. As it reads now, there are a few aphoristic truisms here (e.g., “Causality is dynamic”...”the interplay of religion and mental health depends on broader cultural, political...”) that aren’t helping us get to the point. Too much is going on in this paragraph for the relatively mundane point that things can be dynamic.

- We have rewritten the introduction.

Causal inference and models: 
Paragraphs 3-5 that describe the DAGs need a little attention. It’s a little misleading (perhaps more so to an outsider) to suggest that the MARP data can’t address the backdoor path in Fig. 1b; all it would take is thinking of a variable that would block it. If the MARP data (and our imagination) really can’t do it, some qualification is in order. It reads as unsolvable here for no apparent reason. In other words, the commentary is not presenting the virtues of DAGs very well. It’s tough to distill this large literature into a commentary, but the authors should be able to make things a little clearer, especially since their motivation is to convince readers that causal inference really is the way to go here.

- We think the referee wants us to clarify three points. First, if all common causes (or descendents) of the both the exposure and outcome were to be included in our regression analysis, then it would be possible to obtain an unbiased causal estimate conditional on our causal model. Second, it is likely that all such common causes (or descendents) are captured in the MARP dataset. Third, it is therefore important to suggest why cuasal approach is nevertheless interesting.  We have stated these points clearly in the revised text and furthermore offered a sensitivity analysis to evaluate how large such unmeasured counfounding would need to be to render the causal estimate unreliable at the 95th HPD interval.

As mentioned, Figure 1 is simply unreadable. Even if each panel were larger and more legible, the DAGs are quite messy and the key on the side of the graphs is mysterious; we have 3 circles, one square, and three colors/tones. What do they mean and how are they helping us?

- We agree. We have moved all but the essential

Why are age, SES, education, sex, etc. included here? Yes, there is a large literature that predicts their important (cite them), but since they aren’t discussed here, it seems like an unnecessary and untheorized set of variables that would only confound the estimates even more, given the model.

- First, we have included a reference to VANDERWHEELE 2020 which provides advice on confounder control using these indicators. Second, we note that these indicators are unlikely to measure common effects of the exposures and outcomes; as such inclusion of these indicators is unlikely to introduce collider bias. Third, we note that these indicators are unlikely to stand along the path between the exposures and the outcomes; as such, inclusion of these indicators in our regression models is unlikely to induce mediator bias (or pipe) confounding. Fourth, these indicators are unlikely to capture all potential sources of bias. For example, Vanderwheele 2020 includes a much wider set of potential confounders for confounding control. For this reason, we include a sensitivity analysis, which allows us to state how much coufounding would be necessary to render our causal estimate. The new paragraph reads:

>

The y-axis labels in figure 2 are unreadable. The country-level plots probably aren’t necessary anyway, since that’s not really the focus here.

- Thanks, we agree. We have removed these plots.

Statistical model:   
Justifying the priors here might be useful.

- Thanks, we have included posterior predicted values in the

Appendices: 
code for simulations mod2 and mod3 are identical

- Thanks, fixed.

caption in Figure C1 is probably too preachy for a figure; the figure does not clearly convey the idea that “Causal assumptions are needed for causal inference”. The paper needs a bit more pedagogical flair in text to bring readers to appreciate this point.

- Thanks, figure text fixed.

Details: 
footnote 1: powers[.] Perhaps cite Wallace (1966) or Jensen (2019) for this definition? It’s helpful to be clear about this, but there are many out there (and plenty of folks who have thought hard about this).

- Thanks, we revised the introduction, the citation is no longer needed.

line 40 pg. 3: “makes a start”. colloquial.

-  Thanks, we revised the introduction, no longer relevant.

line 25 pg. 4: “Recall that...” I missed the introduction to this point.

- Thanks, we revised the introduction, no longer relevant.

line 12-13 pg. 5: “these interpretations, we...” something’s not right here.

- Thanks, we revised the introduction, sentence no longer exists.

line 38 pg. 5: “being drawn to God when anxious serves functions”. I get what this sentence means, but this construction is a little awkward. Perhaps rephrase.

- Thanks, agree; again, we revised the introduction. Sentence no longer exists.